INGRID: Creating Languages in MPS from ANTLR Grammars

### 2016-11-12 paper.tex
references
polishing

% coloring and font style commands
  % ANTLR
\newcommand{\antlrap}{\textquotesingle}
\newcommand{\antlrparserrule}[1]{\textcolor{Blue}{#1}}
\newcommand{\antlrlexerrule}[1]{\textcolor{RedOrange}{#1}}
\newcommand{\antlrliteralnoap}[1]{\textcolor{Maroon}{#1}}
\newcommand{\antlrliteral}[1]{\antlrap\antlrliteralnoap{#1}\antlrap}
\newcommand{\antlrregex}[1]{\textcolor{PineGreen}{#1}}
  % MPS in general
\newcommand{\mpsconcept}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\mpsinterface}[1]{\textcolor{Plum}{#1}}
  % MPS structure
\newcommand{\mpsstkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpsstalias}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpsstproperty}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpsstplaceholder}[1]{\textcolor{Gray}{\textbf{#1}}}
\newcommand{\mpsstcardinality}[1]{\textcolor{RoyalBlue}{#1}}
  % MPS editor
\newcommand{\mpsedkeyword}[1]{\textcolor{Black}{\textbf{#1}}}
\newcommand{\mpsedtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpsedannotation}[1]{\textcolor{Gray}{\textbf{#1}}}
  % MPS textgen
\newcommand{\mpstgkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpstgtarget}[1]{\textcolor{Fuchsia}{#1}}
\newcommand{\mpstgparam}[1]{\textcolor{Violet}{#1}}
\newcommand{\mpstgaction}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpstgliteral}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpstgnodeprop}[1]{\textcolor{darkgray}{#1}}

@@

% TODO upravit affiliation pro JetBrains -> doplnit "s.r.o" nebo radsi "Inc."
% TODO muzeme taky napsat jejich emaily pokud budou lidi z JetBrains souhlasit
\authorinfo{V\'aclav Pech}
	{JetBrains}

@@

We present \textsc{Ingrid}, a new method for construction of a language definition in MPS based on its ANTLR grammar.
The structure of a language is generated automatically, but its other aspects --- including the projectional editor --- must be adjusted manually by the user.
During the development of \textsc{Ingrid}, we encountered several challenges related to the mapping between (i) grammars written in an EBNF-like notation and (ii) the language definitions in MPS based on object-oriented principles.
Another difficulty is that grammars do not hold any information about code layout.

We implemented \textsc{Ingrid} as a plugin for MPS and evaluated it on several full-fledged programming languages, such as JavaScript and C\#.
Results show that our approach is practical.
Even though it is only partially automated, it saves the user from many hours of tedious and error-prone work.
Necessary manual adjustments take only a short time.
\end{abstract}

@@

\acks
% TODO
This research was supported by JetBrains. \todo{Pridat jeste nejaky grant a kompletne preformulovat.}

@@

@@

>>>>

% coloring and font style commands
  % ANTLR
\newcommand{\antlrap}{\textquotesingle}
\newcommand{\antlrparserrule}[1]{\textcolor{Blue}{#1}}
\newcommand{\antlrlexerrule}[1]{\textcolor{BrickRed}{#1}}
\newcommand{\antlrliteralnoap}[1]{\textcolor{Maroon}{#1}}
\newcommand{\antlrliteral}[1]{\antlrap\antlrliteralnoap{#1}\antlrap}
\newcommand{\antlrregex}[1]{\textcolor{OliveGreen}{#1}}
  % MPS in general
\newcommand{\mpsconcept}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\mpsinterface}[1]{\textcolor{Plum}{#1}}
  % MPS structure
\newcommand{\mpsstkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpsstalias}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpsstproperty}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpsstplaceholder}[1]{\textcolor{Gray}{\textbf{#1}}}
\newcommand{\mpsstcardinality}[1]{\textcolor{RoyalBlue}{\textbf{#1}}}
  % MPS editor
\newcommand{\mpsedkeyword}[1]{\textcolor{Black}{\textbf{#1}}}
\newcommand{\mpsedtarget}[1]{\textcolor{RubineRed}{\textbf{#1}}}
\newcommand{\mpsedannotation}[1]{\textcolor{Gray}{\textbf{#1}}}
  % MPS textgen
\newcommand{\mpstgkeyword}[1]{\textcolor{Blue}{\textbf{#1}}}
\newcommand{\mpstgtarget}[1]{\textcolor{RubineRed}{\textbf{#1}}}
\newcommand{\mpstgparam}[1]{\textcolor{RoyalPurple}{\textbf{#1}}}
\newcommand{\mpstgaction}[1]{\textcolor{Fuchsia}{\textbf{#1}}}
\newcommand{\mpstgliteral}[1]{\textcolor{Green}{\textbf{#1}}}
\newcommand{\mpstgnodeprop}[1]{\textcolor{darkgray}{\textbf{#1}}}

@@

\authorinfo{V\'aclav Pech}
	{JetBrains}
	{vaclav@jetbrains.com}

@@

We present \textsc{Ingrid}, a method for construction of a language definition in MPS based on its ANTLR grammar.
The structure of a language is generated automatically, but its projectional editor and other aspects must be adjusted manually by the user.
During the development of \textsc{Ingrid}, we encountered several challenges related to the mapping between grammars and language definitions in MPS that are based on object-oriented principles.
Another difficulty is that grammars do not hold any information about code layout.

We implemented \textsc{Ingrid} as a plugin for MPS and evaluated it on several mainstream GPLs, such as JavaScript and C\#.
Results show that our approach is practical, and it saves the user from many hours of tedious and error-prone work.
Necessary manual adjustments take only a short time.
\end{abstract}

@@

\acks
% TODO mozna pridat jeste nejaky dalsi grant (az clanek nekam vezmou)
This work was partially supported by JetBrains.

@@

\bibitem{ref:MPSBOOK} F. Campagne. The MPS Language Workbench, Volume I. CreateSpace Independent Publishing Platform, 2014.

@@

\bibitem{ref:VWK15} M. Voelter, J. Warmer, and B. Kolb. Projecting a Modular Future. IEEE Software, 32(5), 2015.

\bibitem{ref:CSHARP} C\# Language Specification, Version 5.0 (2012), Microsoft Corporation, \url{https://msdn.microsoft.com/en-us/library/ms228593.aspx}
